---
title: "The R package"
---

```{r}
#| echo: false
desc <- read.dcf("../../DESCRIPTION")
pkg <- desc[1, "Package"]
ver <- desc[1, "Version"]
repo_url <- sub(",.*$", "", desc[1, "URL"])
```

`r sprintf("**%s** (v%s)", pkg, ver)` is the computational core of this
dissertation: the methodological R code, its tests, and the Quarto + XeLaTeX
thesis build live together in one package, developed and versioned
independently of the dissertation's web application.

## Installation

```{r}
#| echo: false
#| results: asis
cat(sprintf("```r\n# install.packages(\"remotes\")\nremotes::install_github(\"%s\")\n```\n",
            sub("^https://github.com/", "", grep("github.com", strsplit(desc[1, "URL"], ",\\s*")[[1]], value = TRUE)[1])))
```

Source, issues, and CI live on [GitHub](`r sub("^https://github.com/", "https://github.com/", grep("github.com", strsplit(desc[1, "URL"], ",\\s*")[[1]], value = TRUE)[1])`).

## Function reference

Rendered from the package's roxygen documentation (`man/`) at publish time —
this page cannot drift from the code. Topics are grouped by concept (roxygen
`@family` tags); ungrouped topics fall back to an alphabetical section, so
this page stays correct for any package.

```{r}
#| echo: false
#| results: asis
db <- tools::Rd_db(dir = "../..")

topic_concept <- function(rd) {
  con <- unlist(tools:::.Rd_get_metadata(rd, "concept"))
  if (length(con) == 0) NA_character_ else con[[1]]
}
concepts <- vapply(db, topic_concept, character(1))

emit_topic <- function(nm) {
  tf <- tempfile(fileext = ".html")
  tools::Rd2HTML(db[[nm]], out = tf, fragment = TRUE)
  html <- paste(readLines(tf, warn = FALSE), collapse = "\n")
  cat(sprintf("#### `%s`\n\n", sub("\\.Rd$", "", nm)))
  cat("```{=html}\n<div class=\"rd-topic\">\n", html, "\n</div>\n```\n\n", sep = "")
}

group_order <- c("simulation", "model fitting", "evaluation", "model results")
group_intro <- c(
  simulation = "Seeded, provenance-carrying data generators.",
  `model fitting` = "The model ladder's fitting engines -- conventional baselines, the IRTc product-copula wrapper, and one-factor copulas.",
  evaluation = "Held-out prediction at fixed parameters (the benchmark primary metric).",
  `model results` = "The standardized model-result schema every fit flows through."
)
groups <- unique(c(intersect(group_order, concepts), setdiff(sort(unique(na.omit(concepts))), group_order)))

for (g in groups) {
  cat(sprintf("### %s\n\n", tools::toTitleCase(g)))
  if (!is.na(group_intro[g])) cat(group_intro[[g]], "\n\n")
  for (nm in sort(names(db)[!is.na(concepts) & concepts == g])) emit_topic(nm)
}
ungrouped <- sort(names(db)[is.na(concepts)])
if (length(ungrouped) > 0) {
  cat("### Other topics\n\n")
  for (nm in ungrouped) emit_topic(nm)
}
```
