---
title: "The R package"
---

```{r}
#| echo: false
desc <- read.dcf("../../DESCRIPTION")
pkg <- desc[1, "Package"]
ver <- desc[1, "Version"]
repo_url <- sub(",.*$", "", desc[1, "URL"])
```

`r sprintf("**%s** (v%s)", pkg, ver)` is the computational core of this
dissertation: the methodological R code, its tests, and the Quarto + XeLaTeX
thesis build live together in one package, developed and versioned
independently of the dissertation's web application.

## Installation

```{r}
#| echo: false
#| results: asis
cat(sprintf("```r\n# install.packages(\"remotes\")\nremotes::install_github(\"%s\")\n```\n",
            sub("^https://github.com/", "", grep("github.com", strsplit(desc[1, "URL"], ",\\s*")[[1]], value = TRUE)[1])))
```

Source, issues, and CI live on [GitHub](`r sub("^https://github.com/", "https://github.com/", grep("github.com", strsplit(desc[1, "URL"], ",\\s*")[[1]], value = TRUE)[1])`).

## Function reference

Rendered from the package's roxygen documentation (`man/`) at publish time —
this page cannot drift from the code. Topics are grouped by concept (roxygen
`@family` tags); ungrouped topics fall back to an alphabetical section, so
this page stays correct for any package.

```{r}
#| echo: false
#| results: asis
db <- tools::Rd_db(dir = "../..")

topic_concept <- function(rd) {
  con <- unlist(tools:::.Rd_get_metadata(rd, "concept"))
  if (length(con) == 0) NA_character_ else con[[1]]
}
concepts <- vapply(db, topic_concept, character(1))

# Rd2HTML fragments carry their own <h3> section headings (Title, Usage,
# Arguments, ...) and render arguments as layout tables. Injected raw under an
# h4 function heading, a function's sections outranked the function itself and
# the heading list became "Title, Name, Usage, ..." repeated per topic
# (WCAG 1.3.1 / 2.4.10, wi-20260905-at-rd-reference-structure). Restructure the
# parsed fragment -- no regex over R's HTML:
#   - Title becomes a lead sentence and Name is dropped (the h4 already says it);
#   - remaining sections become h5, each carrying the function name for screen
#     readers, so heading navigation says which function a section belongs to;
#   - argument/value layout tables become description lists (name -> meaning);
#   - argument anchor ids are namespaced per topic (they repeat across topics).
restructure_rd <- function(html, topic) {
  doc <- xml2::read_html(paste0("<div class=\"rd-topic\">", html, "</div>"))
  root <- xml2::xml_find_first(doc, "//div[@class='rd-topic']")

  for (h in xml2::xml_find_all(root, "./h3")) {
    label <- trimws(xml2::xml_text(h))
    if (label == "Name") {
      # The Name section is a bare text node naming the topic -- redundant.
      nxt <- xml2::xml_find_first(h, "following-sibling::node()[normalize-space()][1]")
      if (inherits(nxt, "xml_node") && xml2::xml_type(nxt) == "text") xml2::xml_remove(nxt)
      xml2::xml_remove(h)
    } else if (label == "Title") {
      lead <- xml2::xml_find_first(h, "following-sibling::p[1]")
      if (inherits(lead, "xml_node")) xml2::xml_set_attr(lead, "class", "rd-title lead")
      xml2::xml_remove(h)
    } else {
      xml2::xml_name(h) <- "h5"
      xml2::xml_set_text(h, "")
      prefix <- xml2::xml_add_child(h, "span", paste0(topic, ": "))
      xml2::xml_set_attr(prefix, "class", "visually-hidden")
      xml2::xml_add_child(h, "span", label)
    }
  }

  for (tbl in xml2::xml_find_all(root, ".//table[@role='presentation']")) {
    rows <- xml2::xml_find_all(tbl, ".//tr")
    cells <- lapply(rows, function(tr) xml2::xml_find_all(tr, "./td"))
    if (length(rows) == 0 || !all(lengths(cells) == 2)) next
    dl <- xml2::xml_add_sibling(tbl, "dl", .where = "before")
    xml2::xml_set_attr(dl, "class", "rd-arguments")
    for (pair in cells) {
      dt <- xml2::xml_add_child(dl, "dt")
      for (child in xml2::xml_contents(pair[[1]])) xml2::xml_add_child(dt, child)
      dd <- xml2::xml_add_child(dl, "dd")
      for (child in xml2::xml_contents(pair[[2]])) xml2::xml_add_child(dd, child)
    }
    xml2::xml_remove(tbl)
  }

  for (el in xml2::xml_find_all(root, ".//code[@id]")) {
    xml2::xml_set_attr(el, "id", paste0(topic, "-arg-", xml2::xml_attr(el, "id")))
  }

  paste(vapply(xml2::xml_contents(root), as.character, character(1)), collapse = "")
}

emit_topic <- function(nm) {
  tf <- tempfile(fileext = ".html")
  tools::Rd2HTML(db[[nm]], out = tf, fragment = TRUE)
  html <- paste(readLines(tf, warn = FALSE), collapse = "\n")
  topic <- sub("\\.Rd$", "", nm)
  cat(sprintf("#### `%s`\n\n", topic))
  cat("```{=html}\n<div class=\"rd-topic\">\n", restructure_rd(html, topic), "\n</div>\n```\n\n", sep = "")
}

group_order <- c("simulation", "model fitting", "evaluation", "model results")
group_intro <- c(
  simulation = "Seeded, provenance-carrying data generators.",
  `model fitting` = "The model ladder's fitting engines -- conventional baselines, the IRTc product-copula wrapper, and one-factor copulas.",
  evaluation = "Held-out prediction at fixed parameters (the benchmark primary metric).",
  `model results` = "The standardized model-result schema every fit flows through."
)
groups <- unique(c(intersect(group_order, concepts), setdiff(sort(unique(na.omit(concepts))), group_order)))

for (g in groups) {
  cat(sprintf("### %s\n\n", tools::toTitleCase(g)))
  if (!is.na(group_intro[g])) cat(group_intro[[g]], "\n\n")
  for (nm in sort(names(db)[!is.na(concepts) & concepts == g])) emit_topic(nm)
}
ungrouped <- sort(names(db)[is.na(concepts)])
if (length(ungrouped) > 0) {
  cat("### Other topics\n\n")
  for (nm in ungrouped) emit_topic(nm)
}
```
