---
title: "The R package"
---

```{r}
#| echo: false
desc <- read.dcf("../../DESCRIPTION")
pkg <- desc[1, "Package"]
ver <- desc[1, "Version"]
repo_url <- sub(",.*$", "", desc[1, "URL"])
```

`r sprintf("**%s** (v%s)", pkg, ver)` is the computational core of this
dissertation: the methodological R code, its tests, and the Quarto + XeLaTeX
thesis build live together in one package, developed and versioned
independently of the dissertation's web application.

## Installation

```{r}
#| echo: false
#| results: asis
cat(sprintf("```r\n# install.packages(\"remotes\")\nremotes::install_github(\"%s\")\n```\n",
            sub("^https://github.com/", "", grep("github.com", strsplit(desc[1, "URL"], ",\\s*")[[1]], value = TRUE)[1])))
```

Source, issues, and CI live on [GitHub](`r sub("^https://github.com/", "https://github.com/", grep("github.com", strsplit(desc[1, "URL"], ",\\s*")[[1]], value = TRUE)[1])`).

## Function reference

Rendered from the package's roxygen documentation (`man/`) at publish time —
this page cannot drift from the code.

```{r}
#| echo: false
#| results: asis
db <- tools::Rd_db(dir = "../..")
topics <- sort(names(db))
for (nm in topics) {
  rd <- db[[nm]]
  tf <- tempfile(fileext = ".html")
  tools::Rd2HTML(rd, out = tf, fragment = TRUE)
  html <- paste(readLines(tf, warn = FALSE), collapse = "\n")
  cat(sprintf("### `%s`\n\n", sub("\\.Rd$", "", nm)))
  cat("```{=html}\n<div class=\"rd-topic\">\n", html, "\n</div>\n```\n\n", sep = "")
}
```
